Are Solar Panels Commodities? Evidence of Quality Differences and Asymmetric Information from California

### 2018-06-01 solar_prod_220518.tex
final mcmc version

% solar_prod.tex

\documentclass[12pt]{article}
\usepackage{setspace}
\usepackage[margin=1in]{geometry}
\usepackage{mathtools}
\usepackage{natbib} %for citet and citep
\usepackage{syntonly}
\usepackage{esdiff} %for writing partial derivatives
\usepackage{url} %for inserting urls
\usepackage{placeins}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}

\doublespacing % from package setspacs

\title{Are Solar Panels Commodities? Evidence of Quality Differences and Asymmetric Information from California}

\date{\today}

\author{Johannes Mauritzen \\ BI Norwegian Business School \\ johannes.mauritzen@bi.no\\\url{http://jmaurit.github.io}}

\begin{document}
 \begin{spacing}{1} %sets spacing to single for title page
	\maketitle

\begin{abstract}
 Solar panels should not be considered commodities. Considerable quality differences, as measured directly by degradation of production over time, are found between manufacturers. This has implications for pricing and competition in the market for solar panel systems. I test two implications from the the theory of asymmetric information of quality and find: 1.) Solar systems with high-information third party owners display higher quality than those owned by low-information hosts. 2.) Furthermore, prices of solar panels are more correlated to quality for solar panels that are owned by high-information owners. I use linear mixed effects models estimated by maximum likelihood and multilevel models estimated by Bayesian Markov Chain Monte-Carlo.
\end{abstract}

% \thanks{*I would like to thank...}
% JEL Codes: Q4, L71
 \end{spacing}

@@

Solar panels have been referred to colloquially as being commodities. This is in reference to the high degree of competition, especially from Chinese manufacturers, that has dramatically driven down prices of solar panels in the last decade. This has made solar power price competitive without subsidies in many sunny locations. But it has also meant upheaval in the panel manufacturing industry with cheaper Chinese produced panels replacing panels produced in Europe, North America and Japan. Most recently, this led to the imposition of trade tariffs on Chinese produced solar panels by the US government.

More formally, a commodity is generally defined by a certain degree of fungibility. That is, the good from a certain producer will be largely interchangeable with the good of another producer. Brent crude oil produced by Statoil from the Norwegian North Sea and that produced by BP from the UK North Sea is sold at the same price, as established by an exchange. In a similar manner, the electricity produced by a solar panel at a certain location and time can also be considered a commodity.

@@

The market for rooftop solar panels can be expected to be particularly vulnerable to issues of asymmetric information on quality. Solar panels can be characterised as an ``experience'' good, where an investor/consumer needs to learn about the quality through use. In particular, poor quality panels will tend to show a higher degradation of output over time than high quality panels. Even then, solar panel owners may find it difficult to measure the degradation as it can happen gradually, over many years.

@@

Over the period from 2010 to 2016, many of the larger solar contractors moved to leasing solar power systems to homeowners. The introduction of a leasing model could potentially get over this information asymmetry problem by shifting ownership to the large contractors that install and finance the solar systems \citep{mauritzen_cost_2017}. These contractors can then in turn take steps such as testing panels and visiting manufacturing sites to ensure quality of suppliers. In other words, a descriptive understanding of the market suggests that it is possible to divide the market into two types of solar panel owners: high-information owners who lease the solar panel systems to individual home owners, businesses and other organisations, and low-information owners who buy and own the solar panel systems directly.

@@

In this paper, I use a data set of California solar power systems that includes monthly production data from over 3000 solar panel systems. Where many studies of asymmetric information of quality must use proxies for quality, I measure it directly as the average degradation of production over time.

In summary, I have three nested hypothesis that I wish to test:

\begin{enumerate}
  \item Solar panels are not pure commodities. Significant quality differences can be detected between manufacturers of panels.
  \item High information owners (third-party-owners) will purchase higher quality panels than low-information owners (host-owners)
  \item Price and quality will be more highly correlated in solar panel systems owned by high information types.
\end{enumerate}

The data has a natural hierarchy of between 36 and 60 monthly observations within the 3000-plus solar panel systems, which in turn can be grouped according to relevant characteristics such as whether they are host-owned or third-party owned. Because these groupings display considerable heterogeneity in sample size, mean values and variance, it is natural to model them directly. For relatively parsimonious models with one or two hierarchies and few covariates, I can use linear mixed effects models (LMEM) estimated by maximum likelihood. However, for richer models  with all relevant hierarchies, maximum likelihood becomes impractical.

As a solution, I use a Bayesian multilevel\footnote{Sometimes also called hierarchical} model estimated using Markov Chain Monte Carlo (MCMC) simulation techniques. The main benefit of using a Bayesian multilevel model is that I can fit a complex model with many hundreds of parameters, while avoiding problems of over-fitting and issues of multiple comparison in inference by relying on regularisation of parameters through the use of higher level parameters.

The issue of how information asymmetries of quality affects market outcomes, first introduced by \citet{akerlof_market_1970} is one of the main topics in the modern industrial organisation literature. Ch. 2 of \citet{tirole_theory_1988} and accompanying citation list provides a good overview of the theoretical foundations of this topic.

Early theoretical work of particular relevance to this paper includes \citet{chan_prices_1982}, whom model the situation where information on quality is costly rather than directly unavailable and where quality is endogenous to the seller -- in other words the seller can choose the level of quality of its product. Under these assumptions, the authors predict several equilibrium -- essentially markets for both high and low quality products depending on the extent to which the buyer is informed.

Reputation, or lack thereof, plays a particularly important role in this market. Most of both the solar installers and panel manufacturers are relatively new firms, especially the increasingly dominant Chinese solar panel producers. \citet{shapiro_consumer_1982} explores the role of reputation, and finds that because reputation is effective only with a lag, the "steady-state" quality level chosen by a seller should be lower than in the scenario with perfect information.

An expanding empirical literatures exist on asymmetric information of quality, especially in the marketing, finance and accounting fields. In accounting, the issue of firm quality and voluntary disclosure in regulatory filings as a signal has received broad attention by empirical researchers. \citet{healy_information_2001} review the early literature. Empirical studies in finance have, for example, looked at issues of adverse-selection and quality in Initial Public Offerings \citep{michaely_pricing_1994} and small business lending \citep{petersen_benefits_1994} among many other topics.

The literature in the marketing field may be more relevant to this paper, as it is particularly concerned with the effects of information asymmetry on purchases of consumer durables. The effects of signalling on quality, usually in the form of advertising, is particularly important in the field. \citet{kirmani_no_2000} provide an overview of this literature.

The use of leasing, discussed in this article, can be seen as a form of costly signalling about quality, however leasing plays more than just an informational role since it also transfers much of the risk of poor quality to the contractor. Traditionally, energy generation has been an activity undertaken by large, specialised firms with specialised knowledge and resources. The role of information asymmetry in investment and purchase decisions has likely been seen as a secondary issue, and to my knowledge, no literature exists on the role of such informational issues in the industry.

However, the growth of distributed energy technologies, like solar panels, has added new focus on informational and behavioural issues related to energy investments by homeowners and small businesses. In turn, a growing literature is forming around solar power investment behaviour of homeowners and small businesses.

For example \citet{dastrup_understanding_2012} argue that solar panels can not be considered a pure investment good, but are also bundled as a type of green conspicuous consumption - showing evidence for a ``solar price premium'' in homes with solar panels. \citet{bollinger_peer_2012} study the the role of peer effects in solar power adoption. They find evidence that the adoption of solar panels by homeowners in a certain zip-code will increase the probability that other households in that zip-code will install solar panels.

@@

Table \ref{tbl:variables} gives an overview of the variables in the dataset that are used in the analysis. Notice that variables can vary by monthly observation or by solar power system. I will take account of this natural data hierarchy in my modelling. There are a total of 171,027 monthly observations used in the analysis. These observations come from 3,145 unique solar power systems. From these solar panel systems, panels from a total of 107 unique manufacturers are used.

\begin{table}
  \begin{tabular}{rllll}
  \toprule
    Variable &  Label &   Level & Values &   Mean Value  \\
    \midrule
    Production (kWh) &  prod\_kwh &  Observation & Real & 35,908   \\
    Months of operation &  months\_operation &  Observation & Integer & --   \\
    Month of year & month & Observeration & Categ.$^{1}$& -- \\
    Cost per kw & cost\_per\_kw & System & Real & 7,026 \\
    Capacity rating (kW) & csi\_rating & System & Real & 239 \\
    First year of production & first\_prod\_year & System & Integer & 2011 \\
    Third party owned & lease & System & Categ.$^{2}$  &  35\% leased \\
    Owner sector & own\_sector\_id & System & Categ.$^{3}$& -- \\
    \midrule
    \# Observations && 171,027 && \\
    \# Solar panel systems && 3,145 && \\
    \# Module manufacturers && 107 && \\
     Min. months of data && 36 && \\
     Max. months of data && 60 && \\
  \bottomrule
    \multicolumn{5}{l}{\scriptsize{$^{1}$January - December. $^{2}$Leased 1, Owned 0.$^{3}$ Residential (497), Commercial (1759), Governmental (777), Non-profit (112).}}
  \end{tabular}
  \label{tbl:variables}
  \caption{}
\end{table}

Production is measured in kilowatt hours (kWh). In the analysis, production data is first log transformed and then descaled, as shown in equation \ref{eqn:descaled}. For production data for each system, $s$, the log production data is demeaned and then divided by 2 times the standard deviation of the production data for that solar power system. Figure \ref{fig:production} gives a visualisation of the transformed production data. The red line shows the average degradation over time of all the systems. The strong seasonality of production is also clearly visible.

Other continuous variables are transformed in a similar manner, with observation level variables descaled at the system level, and system-level variables descaled across systems. These transformations avoid confusion that may come when interpreting coefficients on variables with different units and scales. There are also technical reasons for the transformations relating to the MCMC routine. Conversion in the MCMC algorithm is aided when all variables are on the same scale. Setting appropriate weakly informative prior distributions on the parameters is also simplified when continuous variables are free from units.

\begin{equation}
prod\_scaled_{s[i]} = \frac{log(prod\_kwh_{s[i])} - mean(log(prod\_kwh_{s[i]}))}{2 std(log(prod\_kwh_{s[i]}))}
\label{eqn:descaled}
\end{equation}

@@

Month of year dummy variables are used to control for the seasonality in the data. The capacity of the solar panel system is the actual maximum capacity of a solar power system as reported to and audited by the CSI. This can be somewhat lower than the total nameplate capacity of the solar panels, and can depend on not only the technical capacity of the panels, but also on supporting equipment, such as the inverter, and factors such as the angle of the roof and orientation of the solar panel system.

@@

\subsection{Testing for differences between module manufacturer}
Following the notation of \citet{bates_fitting_2015}, the conditional distribution of the response variables, $y$ in a multilevel model can be written compactly as in equation \ref{equation:lme_simple}. Here $\mathbf{\beta}$ is a p-dimensional coefficient vector (or "fixed effects") of the $n times p$ matrix of (p-1) explanatory variables $\mathbf{X}$ plus intercept term. $\mathbf{Z}$ is an $n \times q$ model matrix of the variables whose groups are to be modelled directly with corresponding vector of coefficients, or "random effects" $\mathbf{B}$ that are fixed at values $\mathbf{b}$. \footnote{The use of "random" refers to the fact that these groups are assumed to be specific to the sample. Many applications of mixed effects model thus treat the varying group coefficients as group-level error-terms, thus the "random" label. Still, interpreting coefficients as group varying or as group error terms is mathematically equivalent.} Here $q$ is a dependent on the number of terms, $k$, that are allowed to vary by group and the number of groups, $l$ for each term: $q=\sum_i^k q_i = \sum_i^k l_i p_i$.  $\mathbf{W}$ is a diagonal matrix of known prior weights. $B$ is distributed as multivariable normal with zero mean and where $\Sigma$ is variance-covariance matrix.

@@

Models of this form can be solved efficiently using a penalised maximum likelihood routine. For further details of both the model and routine, I refer to \citet{bates_fitting_2015}.

@@

In this model, the transformed production variable, $prod\_scaled_{i}$ for observation $i$ is distributed normally with the mean modelled as a linear function of an intercept term $a_s$, a slope term, $b_s$ on the transformed months of operation variable $months\_operation\_scaled$, and a set of dummy variables representing the month of year, $\mathbf{month}$. Notice that the intercept and slope are indexed by $s$, indicating that they are allowed to vary by the individual system. Thus in this model formulation I take into account the groupings by system, explicitly modelling varying relative initial production values (intercepts) and degradation over time (slope). Moving up a level, the estimated $a$ and $b$ parameters are in turn modelled as multivariate normal.

\begin{equation}
\begin{aligned}
prod\_scaled_{i} &\sim N(\mathbf{month} + a_s + b_s months\_operation\_scaled_{i}, \sigma^2)\\
\begin{pmatrix}
  a\\
  b
\end{pmatrix}
&\sim N
\begin{pmatrix}
  \mu_a\\
  \mu_b
\end{pmatrix},
\begin{pmatrix}
  \sigma_a^2 & \rho \sigma_a \sigma_b \\
  \rho \sigma_a \sigma_b & \sigma_b^2
\end{pmatrix}\label{eqn:observation_level_rest}
\end{aligned}
\end{equation}

The inclusion of the matrix of month dummies $\mathbf{month}$ controls for seasonal variation across all systems. This is a simplified way to model seasonality, but because the data is itself at a monthly frequency, and the seasonality is not of direct importance, this should suffice. As long as un-modeled variation in seasonality is not correlated with the long-run slope of production, the point estimates for the slope should be unbiased.

To test whether quality varies among module manufacturers, I simply add panel manufacturer as an extra hierarchy to the model. In this model, the observation level regression is identical to the restricted comparison model as in equation \ref{eqn:observation_level_rest}. The system-level intercepts, $a$, and slope, $b$, terms are also modelled as a multivariate normal. But the mean values of the multivariate normal distribution, $\mu_a$ and $\mu_b$, are now indexed by $m$, as shown in equation \ref{eqn:unrest}, representing that they are allowed to vary by module manufacturer.

\begin{equation}
\begin{aligned}
\begin{pmatrix}
  a\\
  b
\end{pmatrix}
&\sim N
\begin{pmatrix}
  \mu_{a[m]}\\
  \mu_{b[m]}
\end{pmatrix},
\begin{pmatrix}
  \sigma_{a[m]}^2 & \rho \sigma_{a[m]} \sigma_{b[m]} \\
  \rho \sigma_a \sigma_b & \sigma_b^2
\end{pmatrix} \\
\mu_{a} &\sim N(\theta_a, \sigma_{\mu_a}^2) \\
\mu_{b} & \sim N(\theta_b, \sigma_{\mu_b}^2) \label{eqn:unrest}
\end{aligned}
\end{equation}

If there is no significant variation between the module manufacturer groupings, this model collapses into the restricted comparison model, which is the null hypothesis. Figure \ref{sfig:sys_slope_fig} gives a visual summary of the system-level slope terms ordered from lowest to highest. The lines represent approximate 95\% confidence intervals around the point estimate. One anomaly we can notice is that a handful of the systems have slopes that are estimated to be positive. This is clearly problematic if we wish to interpret the slope coefficient as degradation over time. I address this in the full Bayesian model without resorting to excluding data.

\begin{figure}
\begin{minipage}{.48\textwidth}
  \includegraphics[width=1\linewidth]{figures/sys_slope_fig.png}
  \caption{Approximate 95\% confidence intervals around maximum likelihood point estimates of system-level slope terms, ordered from lowest to highest.}
  \label{sfig:sys_slope_fig}
\end{minipage}\hfill
\begin{minipage}{.48\textwidth}
  \includegraphics[width=1\linewidth]{figures/nested_manuf_fig_final.png}
  \caption{Approximate 95\% confidence intervals around maximum likelihood estimators of manufacturer-level slope terms $\mu_b$, ordered from lowest to highest.}
  \label{sfig:nested_manuf_fig}
\end{minipage}
\end{figure}

\ref{sfig:nested_manuf_fig} gives a summary of the estimated manufacturer-level mean slopes. Again the lines represent 95\% confidence intervals around the point estimates. Visually, we can see substantial variation in quality between the manufacturer groups. We can more formally test the hypothesis that manufacturer groupings improve the explanatory power of the model by comparing to the restricted model with an LM test. The results are shown in table \ref{tbl:lm_commodity}. The information criteria measures (AIC, BIC) are both lower, indicating a better predictive power of the model with manufacturer groupings. The chis-square statistic also indicates that the model with manufacturer groupings substantially improves model performance.

>>>>

\documentclass[a4paper]{article}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=black]{hyperref}
\usepackage{natbib}
\usepackage{url}
\usepackage{placeins} %for float barrier

%spacing
\usepackage{setspace}
\doublespacing
%\onehalfspacing

\title{Are Solar Panels Commodities? Evidence of Quality Differences and Asymmetric Information from California}
\author{Johannes Mauritzen\thanks{BI Norwegian Business School, johannes.mauritzen@bi.no, \url{http://jmaurit.github.io}}}

\begin{document}
\maketitle

\begin{abstract}
Solar panels should not be considered commodities. Considerable quality differences, as measured directly by degradation of production over time, are found between manufacturers. This has implications for pricing and competition in the market for solar panel systems. I test two implications from the the theory of asymmetric information of quality and find: 1.) Solar systems with high-information third party owners display higher quality than those owned by low-information hosts. 2.) Furthermore, the price of solar panels that are owned by high-information owners are more highly correlated to quality. I use random effects models estimated by maximum likelihood and hierarchical models estimated by Bayesian Markov Chain Monte-Carlo.
\end{abstract}

\newpage{}

@@

Solar panels have been referred to colloquially as being commodities.\footnote{See for example \url{https://www.greentechmedia.com/articles/read/are-pv-modules-a-commodity\#gs.lx3tMbY}} This is often in reference to the high degree of competition, especially from Chinese manufacturers, that has dramatically driven down prices of solar panels in the last decade. This has made solar power price competitive without subsidies in many sunny locations. But it has also meant upheaval in the panel manufacturing industry with cheaper Chinese produced panels replacing panels produced in Europe, North America and Japan. Most recently, this led to the imposition of trade tariffs on Chinese produced solar panels by the US government.

More formally, a commodity is generally defined by a certain degree of fungibility. That is, the good from a certain producer will be largely interchangeable with the good of another producer. Brent crude oil produced by Statoil from the Norwegian North Sea and that produced by BP from the UK North Sea are sold at the same price, as established by an exchange. In a similar manner, the electricity produced by a solar panel at a certain location and time can also be considered a commodity.

@@

The market for rooftop solar panels can be expected to be particularly vulnerable to issues of asymmetric information on quality. Solar panels can be characterized as an ``experience'' good, where an investor/consumer needs to learn about the quality through use. In particular, poor quality panels will tend to show a higher degradation of output over time than high quality panels. Even then, solar panel owners may find it difficult to measure the degradation as it can happen gradually, over many years.

@@

Over the period from 2010 to 2016, many of the larger solar contractors moved to leasing solar power systems to home- and business-owners \citep{mauritzen_cost_2017}. These contractors, who buy in scale, have the resources and expertise to judge quality in solar panels. They can take steps such as getting panels tested at specialized engineering firms and visiting manufacturing sites to ensure quality of suppliers. This descriptive understanding of the market suggests that it is possible to divide the market into two types of solar panel owners: high-information owners who lease the solar panel systems to individual home owners, businesses and other organizations, and low-information owners who buy and own the solar panel systems directly.

@@

In this paper, I use a data set of California solar power systems that includes monthly production data from over 3000 solar panel systems. Where many studies of asymmetric information of quality use proxies for quality, I measure it directly as the average degradation of production over time.

I have three nested predictions that I wish to test:

\begin{enumerate}
  \item Solar panels are not pure commodities. Significant quality differences can be detected between manufacturers of panels.
  \item Given that quality differences exist between manufacturers, high information owners (third-party owners) will purchase higher quality panels than low-information owners (host owners).
  \item Price and quality will be more highly correlated in solar panel systems owned by high information owners.
\end{enumerate}

The data has a natural hierarchy of between 36 and 60 monthly observations within the 3000-plus solar panel systems, which in turn can be grouped according to relevant characteristics such as whether they are host-owned or third-party owned. Because these groupings display considerable heterogeneity in sample size, mean values and variance, it is natural to model the relevant structure of the data directly. For relatively parsimonious models with one or two hierarchies and few covariates, I can use linear mixed effects models (LMEM) estimated by maximum likelihood. However, for richer models with all relevant hierarchies, maximum likelihood becomes impractical.

As a solution, I use a Bayesian hierarchical\footnote{Sometimes also called multilevel modeling or random effects modeling} model estimated using Markov Chain Monte Carlo (MCMC) simulation techniques. This form of modeling has long been established in multivariate dynamic modeling in macroeconomics \citep{doan_forecasting_1984, litterman_forecasting_1986,sims_bayesian_1998, canova_forecasting_2004},  but is also increasingly common in microeconomics and corporate finance. One of the main advantages is superior out of sample predictive properties with data with distinct groupings. For example \citet{dehejia_was_2003} uses a Bayesian hierarchical model to evaluate multiple programs at different sites. By incorporating ``predictive uncertainty'' by way of a hierarchical model, the impact of the sites, which otherwise is estimated to be significant with a fixed-effects model, disappears. \citet{meager_understanding_2018} presents a more recent application in the microeconomics litterature. She uses a hierarchical model to aggregate studies of microcredit expansion in developing countries in order to estimate an overall treatment effect with ``external validity,'' that is the degree to which an estimated treatment effect generalizes to other populations.

Within the field of corporate finance, \citet{jones_mutual_2005} uses a hierarchical model to study returns of mutual funds, arguing that estimates of long-run returns of any single fund should be informed, by way of a higher-level distribution, of the returns of the population of other mutual funds. In a study with parallels to my own, \citet{korteweg_skill_2017} uses a Bayesian hierarchical model to study persistence of returns in the private equity sector both between and within firms, where persistence is measured by excess variation between firms as captured by firm-level parameters (or random effects) that are drawn from a higher-level distribution. If returns are thought of as the converse of production degradation, then the existence of persistence of returns can be thought of as the converse of quality differences.

%In another early application, \citep{chamberlain_hierarchical_1996} used a Bayesian hierarchical model to aggregate multiple weak instruments.

%Corporate finance: Korteweg_markov_2011, Korteweg_skill_2017

The property of hierarchical models that allows for improved predictive performance with the presence of groupings in the data is regularization, or alternatively called partial pooling \citep{gelman_bayesian_2013}. As opposed to fixed-effects estimation where a fixed-effects coefficient is determined exclusively with data from within a particular group, in a hierarchical model each of the group parameters are regularized towards a higher-level parameter estimated with the full data in proportion to the relative number of data points in a particular group.

The linear mixed effects models estimated by maximum likelihood that I use in the first section of this article also have have the regularization property. Bayesian estimation provides several theoretic advantages, such as correct inference in small samples and generalization to distributions other than the normal. However, for my purproses, the main benefit of using a Bayesian methodology is practical. I can fit a rich model with several hierarchies and many thousands of parameters, that simply can not be handled by maximum likelihood procedures.

The issue of how information asymmetries of quality affects market outcomes, first introduced by \citet{akerlof_market_1970} is one of the main topics in the modern industrial organization literature. Ch. 2 of \citet{tirole_theory_1988} and accompanying citation list provides a good overview of the theoretical foundations of this topic.

Early theoretical work of particular relevance to this paper includes \citet{chan_prices_1982}, whom model the situation where information on quality is costly rather than directly unavailable and where quality is endogenous to the seller--in other words the seller can choose the level of quality of its product. Under these assumptions, the authors predict several equilibrium--essentially markets for both high and low quality products depending on the extent to which the buyer is informed.

Reputation, or lack thereof, plays a particularly important role in this market. Most solar installers and panel manufacturers are relatively new firms, especially the increasingly dominant Chinese solar panel producers. \citet{shapiro_consumer_1982} explores the role of reputation, and finds that because reputation is effective only with a lag, the "steady-state" quality level chosen by a seller should be lower than in the scenario with perfect information.

An expanding empirical literatures exist on asymmetric information of quality, especially in the marketing, finance and accounting fields. In accounting, the issue of firm quality and voluntary disclosure in regulatory filings as a signal has received broad attention by empirical researchers. \citet{healy_information_2001} review the early literature. Empirical studies in finance have, for example, looked at issues of adverse-selection and quality in Initial Public Offerings \citep{michaely_pricing_1994}, small business lending \citep{petersen_benefits_1994}, and in a recent application for consumer credit markets \citep{dobbie_information_2013}.

The literature in the marketing field is also relevant to this paper, as it is particularly concerned with the effects of information asymmetry on purchases of consumer durables. The effects of signaling on quality, usually in the form of advertising, is particularly important in the field. \citet{kirmani_no_2000} provide an overview of this literature.

The use of leasing, discussed in this article, can be seen as a form of costly signaling about quality, however leasing plays more than just an informational role since it also transfers much of the risk of quality to the contractor. Traditionally, energy generation has been an activity undertaken by large, specialized firms with specialized knowledge and resources. The role of information asymmetry in investment and purchase decisions has likely been seen as a secondary issue, and to my knowledge, no literature exists on the role of such informational issues in the industry.

However, the growth of distributed energy technologies, like solar panels, has added new focus on informational and behavioral issues related to energy investments by homeowners and small businesses. In turn, a growing literature is forming around solar power investment behavior of homeowners and small businesses.

For example \citet{dastrup_understanding_2012} argue that solar panels can not be considered a pure investment good, but are also bundled as a type of green conspicuous consumption, showing evidence for a ``solar price premium'' in homes with solar panels. \citet{bollinger_peer_2012} study the the role of peer effects in solar power adoption. They find evidence that the adoption of solar panels by homeowners in a certain zip-code will increase the probability that other households in that zip-code will install solar panels.

@@

Table \ref{tbl:variables} gives an overview of the variables in the dataset that are used in the analysis. Notice that variables can vary by monthly observation or by solar power system. I will take account of this natural data hierarchy in my modeling. There are a total of 171,027 monthly observations used in the analysis. These observations come from 3,145 unique solar power systems. From these solar panel systems, panels from a total of 107 unique manufacturers are used.

\begin{table}
  \caption{Summary of data. Information is provided on whether whether the variables are at the observation (system-month) level or at the system level, Whether the variable values are real, integer or categorical in nature and the mean value of observations is also provided if relevant. The parenthesis after the sector categories in the footnote indicate the number of systems for each sector.}
  \begin{tabular}{rllll}
  \toprule
    Variable &  Label &   Level & Values &   Mean Value  \\
    \midrule
    Production (kWh) &  prod\_kwh &  Observation & Real & 35,908   \\
    Months of operation &  months\_operation &  Observation & Integer & --   \\
    Month of year & month & Observation & Categ.$^{1}$& -- \\
    Cost per kw & cost\_per\_kw & System & Real & 7,026 \\
    Capacity rating (kW) & csi\_rating & System & Real & 239 \\
    First year of production & first\_prod\_year & System & Integer & 2011 \\
    Third party owned & lease & System & Categ.$^{2}$  &  35\% leased \\
    Owner sector & own\_sector\_id & System & Categ.$^{3}$& -- \\
    Host county & county\_id & System & Categ. & -- \\
    \midrule
    \# Observations && 171,027 && \\
    \# Solar panel systems && 3,145 && \\
    \# Module manufacturers && 107 && \\
     Min. months of data && 36 && \\
     Max. months of data && 60 && \\
  \bottomrule
    \multicolumn{5}{l}{\scriptsize{$^{1}$January - December. $^{2}$Leased 1, Owned 0.$^{3}$ Residential (497), Commercial (1759), Governmental (777), Non-profit (112).}}
  \end{tabular}
  \label{tbl:variables}
\end{table}

Production is measured in kilowatt hours (kWh). In the analysis, production data is first log transformed and then de-scaled, as shown in equation \ref{eqn:descaled}. For production data for each system, $s$, the log production data is demeaned and then divided by 2 times the standard deviation of the production data for that solar power system. Figure \ref{fig:production} gives a visualization of the transformed production data. The red line shows the average degradation over time of all the systems. The strong seasonality of production is also clearly visible.

Other continuous variables are transformed in a similar manner, with observation level variables de-scaled at the system level, and system-level variables de-scaled across systems. These transformations avoid confusion that may come when interpreting coefficients on variables with different units and scales. There are also technical reasons for the transformations relating to the MCMC routine. Convergence in the MCMC routine is aided when all variables are on the same scale. Setting appropriate weakly informative prior distributions on the parameters is also simplified when continuous variables are free from units.

\begin{equation}
prod_{s[i]} = \frac{prod\_kwh_{s[i]} - mean(prod\_kwh_{s[i]})}{std(prod\_kwh_{s[i]})}
\label{eqn:descaled}
\end{equation}

@@

Month-of-year dummy variables are used to control for the seasonality in the data. The capacity of the solar panel system is the actual maximum capacity of a solar power system as reported to and audited by the CSI. This can be somewhat lower than the total nameplate capacity of the solar panels, and can depend on not only the technical capacity of the panels, but also on supporting equipment, such as the inverter, and factors such as the angle of the roof and orientation of the solar panel system.

@@

\subsection{Testing for differences between module manufacturer}
Following the notation of \citet{bates_fitting_2015}, the conditional distribution of the response variables, $y$ in a linear mixed effects model can be written compactly as in equation \ref{equation:lme_simple}. Here $\mathbf{\beta}$ is a p-dimensional coefficient vector (or  ``fixed effects'') of the $n \times p$ matrix $\mathbf{X}$ of (p-1) explanatory variables plus intercept term. $\mathbf{Z}$ is an $n \times q$ model matrix of the variables whose groups are to be modeled directly with corresponding vector of coefficients, or ``random effects'' $\mathbf{B}$ that are fixed at values $\mathbf{b}$.\footnote{The use of ``random'' refers to the fact that these groups are assumed to be specific to the sample. Many applications of mixed effects model thus treat the varying group coefficients as group-level error-terms. Interpreting coefficients as group varying coefficients or as group error terms is mathematically equivalent.} Here $q$ is a dependent on the number of terms, $k$, that are allowed to vary by group and the number of groups, $l$ for each term: $q=\sum_i^k q_i = \sum_i^k l_i p_i$.  $\mathbf{W}$ is a diagonal matrix of known prior weights. $B$ is distributed as multinomial normal with zero mean and where $\Sigma$ is variance-covariance matrix.

@@

Models of this form can be solved efficiently using a penalized maximum likelihood routine. For further details of both the model and routine, I refer to \citet{bates_fitting_2015}.

@@

In this model, the transformed production variable, $prod\_scaled_{i}$ for observation $i$ is distributed normally with the mean modeled as a linear function of an intercept term $a_s$, a slope term, $b_s$ on the transformed months of operation variable $months\_operation\_scaled$, and a set of dummy variables representing the month of year, $\mathbf{month}$. Notice that the intercept and slope are indexed by $s$, indicating that they are allowed to vary by the individual system. Thus in this model formulation I take into account the groupings by system, explicitly modeling varying relative initial production values (intercepts) and degradation over time (slope). Moving up a level, the estimated $a$ and $b$ parameters are in turn modeled as multivariate normal.

\begin{equation}
\begin{aligned}
prod_{i} &\sim N(\mathbf{month} + a_s + b_s months\_operation\_scaled_{i}, \sigma^2)\\
\begin{pmatrix}
  a\\
  b
\end{pmatrix}
&\sim N
\begin{pmatrix}
  \mu_a\\
  \mu_b
\end{pmatrix},
\begin{pmatrix}
  \sigma_a^2 & \rho \sigma_a \sigma_b \\
  \rho \sigma_a \sigma_b & \sigma_b^2
\end{pmatrix}\label{eqn:observation_level_rest}
\end{aligned}
\end{equation}

The inclusion of the matrix of month dummies, $\mathbf{\mu^{month}}$, controls for seasonal variation across all systems. This is a simplified way to model seasonality, but because the data is itself at a monthly frequency, and a more direct modeling of seasonality is not of direct importance, this should suffice. As long as un-modeled variation in seasonality is not correlated with the long-run slope of production, the point estimates for the slope should be unbiased.

To test whether quality varies among module manufacturers, I simply add panel manufacturer as an extra hierarchy to the model. In this model, the observation-level regression is identical to the restricted comparison model as in equations \ref{eqn:observation_level_rest}. The system-level intercepts, $a$, and slopes, $b$, terms are also modeled as multivariate normal. But the mean values of the multivariate normal distribution, $\mu_a$ and $\mu_b$, are now indexed by $m$, as shown in equation \ref{eqn:unrest}, representing that they are allowed to vary by module manufacturer.

\begin{equation}
\begin{aligned}
\begin{pmatrix}
  a\\
  b
\end{pmatrix}
&\sim N \Bigg(
\begin{pmatrix}
  \mu_{a[m]}\\
  \mu_{b[m]}
\end{pmatrix},
\begin{pmatrix}
  \sigma_{a[m]}^2 & \rho \sigma_{a} \sigma_{b} \\
  \rho \sigma_a \sigma_b & \sigma_b^2
\end{pmatrix} \Bigg) \\
\mu_{a} &\sim N(\theta_a, \sigma_{\mu_a}^2) \\
\mu_{b} & \sim N(\theta_b, \sigma_{\mu_b}^2) \label{eqn:unrest}
\end{aligned}
\end{equation}

If there is no significant variation between the module manufacturer groupings, this model collapses into the restricted comparison model, which is the null hypothesis. Figure \ref{sfig:sys_slope_fig} gives a visual summary of the system-level slope terms ordered from lowest to highest. The lines represent approximate 95\% confidence intervals around the point estimates. One anomaly we can notice is that a handful of the systems have slopes that are estimated to be positive. This is clearly problematic if we wish to interpret the slope coefficient as degradation over time. I address this in the full Bayesian model without resorting to excluding data.

\begin{figure}
  \caption{Approximate 95\% confidence intervals around maximum likelihood point estimates of system-level slope terms, $b_i$, ordered from lowest to highest. Visually, we can see significant differences in the average degradation over time across solar power systems.}
  \includegraphics[width=1\linewidth]{figures/sys_slope_fig.png}
  \label{sfig:sys_slope_fig}
\end{figure}

Figure \ref{sfig:nested_manuf_fig} gives a summary of the estimated manufacturer-level mean slopes. Again the lines represent 95\% confidence intervals around the point estimates. Visually, we can see substantial variation in quality between the manufacturer groups. We can formally test the hypothesis that manufacturer groupings improve the explanatory power of the model with a comparison to the restricted model with a likelihood ratio test. The results are shown in table \ref{tbl:lm_commodity}. The information criteria measures (AIC, BIC) are both lower, indicating better predictive power of the model with manufacturer groupings. The Chi-square statistic indicates that the model with manufacturer groupings substantially improves model performance.

\begin{figure}
  \caption{Approximate 95\% confidence intervals around maximum likelihood estimators of manufacturer-level mean slope terms $\mu_b$, ordered from lowest to highest. These are the means of the higher level normal distributions that vary by manufacturer that the system-level parameters are assumed to be drawn from. The visual impression is of substantial variation between manufacturer groupings.}
  \includegraphics[width=1\linewidth]{figures/nested_manuf_fig_final.png}
  \label{sfig:nested_manuf_fig}
\end{figure}
